ACNP 2024 MS ANXIETY

### 2024-08-22 scripts/MSAnxiety_manuscript_audrey_replication.Rmd
added IF/Steroids sensitivity analysis

##### Physical and Mental Health x anxiety dose intx
```{r phys_and_mental_x_anxiety_dose}
###
df_for_promis_plot_mental_vs_physical_healthy_no_summarize <- df_people_with_promis_data_scores_added_unique %>%
  group_by(EMPI) %>%
  mutate(physical_health_score = sum(across(physical_domains))/length(physical_domains)) %>%
  mutate(mental_health_score = sum(across(mood_domains))/length(mood_domains)) %>%
  mutate(anxiety_dose = as.numeric(anxiety_dose)) %>%
  dplyr::select(EMPI,
                anxiety_dose,
                physical_health_score,
                mental_health_score) %>%
  gather(health_variable, PROMIS_Score, -anxiety_dose, -EMPI)

@@

@@

### Sex Differences lm and chisq
```{r diseaseXsexdiff}
############## sex diff
#df_anxiety_disorder_by_sex <-

@@

#### healthy vs super sick - values that fdr correct, most not anxiety net
```{r individual_fascicle_disease_healthy_vs_super_sick}
df_demo_and_fascicles_no_unclass_anxiety_dose <- df_demo_and_fascicles %>%
  mutate(anxiousAndMeds_or_healthy = oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed) %>%
  filter(anxiousAndMeds_or_healthy %in% c("true_healthy", "Has.Anxiety.And.On.Anxiety.Meds.Inc.Dep")) %>%
  mutate(anxiousAndMeds_or_healthy = factor(anxiousAndMeds_or_healthy, levels=c("true_healthy", "Has.Anxiety.And.On.Anxiety.Meds.Inc.Dep")))

>>>>

##### Physical and Mental Health x anxiety dose intx -N.S.
```{r phys_and_mental_x_anxiety_dose}
###
df_for_promis_plot_mental_vs_physical_healthy_no_summarize <- df_people_with_promis_data_scores_added_unique %>%
  group_by(EMPI) %>%
  mutate(physical_health_score = sum(across(physical_domains))/length(physical_domains)) %>%
  mutate(mental_health_score = sum(across(mood_domains))/length(mood_domains)) %>%
  mutate(anxiety_dose = as.numeric(anxiety_dose)) %>%
  dplyr::select(EMPI,
                anxiety_dose,
                physical_health_score,
                mental_health_score) %>%
  gather(health_variable, PROMIS_Score, -anxiety_dose, -EMPI)

@@

##### No steroids/IF -N.S. but low n's
```{r no_steroids_or_if}
#### Steroids and interferon are known to have psychiatric side effects
### we repeated our analyses but removed these groups
### Nothing sig but trend, probably because we lose N (2 people in MS+noA, 1 in MS+mildA, 2 in MS+severeA)

### remove people on IF or steroids
df_demo_and_fascicles_no_steroids_or_ifn <- df_demo_and_fascicles %>%
  filter((On.Steroids == 0) & (On.Interferon == 0))

### print demographics, MS+noA = 96, MS+mildA = 248, MS+severeA=21
make_demographics_table_ms_anxiety_simple(data_frame = df_demo_and_fascicles_no_steroids_or_ifn[df_demo_and_fascicles_no_steroids_or_ifn$oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified",])

make_demographics_table_ms_anxiety_simple(data_frame = df_demo_and_fascicles_no_steroids_or_ifn[((df_demo_and_fascicles_no_steroids_or_ifn$oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") & (df_demo_and_fascicles_no_steroids_or_ifn$oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "Has.Anxietydx.OR.Antianxiety.Meds")),])

## Make df for Anxiety Diagnosis
#define group
  uncinate_by_dx_df <- df_demo_and_fascicles_no_steroids_or_ifn %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "Has.Anxietydx.OR.Antianxiety.Meds") %>%
  mutate(subject = as.factor(EMPI)) %>%
  mutate(Diagnosis = oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed) %>%
  mutate(Diagnosis = ifelse(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed=="true_healthy", "MS-Anxiety", "MS+Anxiety")) %>%
  mutate(Diagnosis = as.factor(Diagnosis)) %>%
 # dplyr::select(subject, osex, Diagnosis, PAT_AGE_AT_EXAM, UF_R_vol, UF_L_vol, F_R_vol, F_L_vol, mean_UF_vol) %>%
  ungroup()

#bilateral UF ~ Diagnosis: DiagnosisMS+Anxiety    768.2      422.2   1.820   0.0715 .  trend, likely due to sample size
disease_mean_UF_vol_gam <- gam(mean_UF_vol ~ Diagnosis + osex+s(PAT_AGE_AT_EXAM, k = 4, fx = F), data = uncinate_by_dx_df)
summary(disease_mean_UF_vol_gam)

##test Anxiety Severity Model
#define group
df_demo_and_fascicles_no_unclass_anxiety_dose <- df_demo_and_fascicles_no_steroids_or_ifn %>%
  filter(oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed != "unclassified") %>%
  filter(anxiety_dose != -1) #n=371

  #bilat anxiety dose + sex + age: anxiety_dose: anxiety_dose    369.9      189.7   1.949    0.052 . (trend, likely due to sample size)   
  anxietyGrouping_mean_UF_by_sex_gam <- gam(mean_UF_vol ~ anxiety_dose + osex + s(PAT_AGE_AT_EXAM, k = 4, fx = F), data=df_demo_and_fascicles_no_unclass_anxiety_dose)
  summary(anxietyGrouping_mean_UF_by_sex_gam)

@@

### Sex Differences lm and chisq - N.S.
```{r diseaseXsexdiff}
############## sex diff
#df_anxiety_disorder_by_sex <-

@@

#### healthy vs super sick - values that fdr correct, most not anxiety net
```{r individual_fascicle_disease_healthy_vs_super_sick}

df_demo_and_fascicles_no_unclass_anxiety_dose <- df_demo_and_fascicles %>%
  mutate(anxiousAndMeds_or_healthy = oanxietydx_OR_meds_AND_Anxietydx.Meds.And.Dep.True.Healthy.Collapsed) %>%
  filter(anxiousAndMeds_or_healthy %in% c("true_healthy", "Has.Anxiety.And.On.Anxiety.Meds.Inc.Dep")) %>%
  mutate(anxiousAndMeds_or_healthy = factor(anxiousAndMeds_or_healthy, levels=c("true_healthy", "Has.Anxiety.And.On.Anxiety.Meds.Inc.Dep")))
